---
title: "Machine learning project"
author: "Keilor Rojas"
date: "26 de julio de 2015"
output: html_document
---

The aim of this project is to building model that can determine how well people 
perform barbell lifts, based on data from accelerometers on the belt, forearm, arm, 
and dumbell collected from 6 participants and freely available in the following link
[http://groupware.les.inf.puc-rio.br/har]

##open packages
```{r,results='hide'}
library(caret)
library(ggplot2)
library(VIM)
library(gridExtra)
library(knitr)
library(rpart)
library(rattle)
library(MASS)
library(randomForest)
library(GGally)
library(klaR)
```

##load data
```{r,results='hide'}
set.seed(345)
tr <- read.csv("pml-training.csv")
te <- read.csv("pml-testing.csv")
names(tr)
str(tr)
```

##simplify analysis by removing some noise columns and uncomplete data
```{r}
columns <- "^(kurtosis|skewness|min|max|stddev|total|var|avg|ampl)"
data <- tr[,grep(columns,names(tr),invert=T)]
table(complete.cases(data))
```

##impute the data
```{r,results='hide'}
con <- aggr(tr,plot=F)
summary(con)
```
```{r}
table(con$missings$Count)
data <- data[,grep("^(num_window|cvtd_timestamp|X|new_window)",names(data),invert=T)]
```

##subdivide training set into subtraining and validation set
```{r}
inTraining <- createDataPartition(y=data$classe,p=.75,list=F)
training <- data[inTraining,]
validation <- data[-inTraining,]

tr_pred <- training[,-c(1,52)]
tr_out <- training[,c(52)]

val_pred <- validation[,-c(52)]
val_outcome <- validation[,c(52)]
```

##model1 Regression tree with preprocessing of data and traicontrol
```{r}
trcontrol <- trainControl(method="cv",number=5,allowParallel=T)
model1 <- train(tr_out ~ ., data=tr_pred, preProcess=c("center","scale"),
                method="rpart", trControl=trcontrol)
model1
model1$finalModel
```

```{r, echo=FALSE}
fancyRpartPlot(model1$finalModel)
```

provides a nice visual indication of main predictor variables but accuracy is still low

##model2 Lda
```{r}
model2 <- train(tr_out ~ .,data=tr_pred, preProcess=c("center","scale"),
                method="lda", trControl=trcontrol)
model2
```
lda had a better performance than regression tree model but accuracy is still low

##model3 Random Forest
```{r}
model3 <- train(tr_out ~ .,data=tr_pred, preProcess=c("center","scale"),
                method="rf", prof = TRUE, trControl=trcontrol)
model3
```
this model has a much better performance. The accuracy was 99%


##cross validation with validation subset and expected out of sample error
```{r}
pred <- predict(model3, validation)
confMat <- confusionMatrix(pred, val_outcome)
confMat
```
Random Forest model provided High accuracy and low error

##prediction for 20 cases of test set 
```{r}
pred2 <- predict(model3, te)
pred2
```
this will lead to the predictions of the 20 cases

##Acknowledgment
Data were kindly provided by http://groupware.les.inf.puc-rio.br/har


